---
title: "Lab2-3-Rnash"
author: "Raphael Nash"
date: "September 3, 2016"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
 
```

Your task is to create an R dataframe that shows rates of tuberculosis infection by country. You shouldusetheinformationfromthetbdatabaseandtheattachedpopulation.csvfile. YourR dataframe should have the following columns:
     Country, Year, Rate
Where Rate is defined as Cases/Population.


```{r, warning=FALSE, message = FALSE}

library("RMySQL")
library("DT")
library("knitr")
library("dplyr")
library("ggplot2")
library("plotly")
conn <- dbConnect(MySQL(), user='rnash', host='localhost', dbname='tb')

```

```{r}

population <- read.csv("population.csv")
str(population) 
kable(head(population))

```



```{r warning=FALSE}

tbCases <- dbGetQuery(conn, "select 
                                country, year, 
                                sum( coalesce(child, 0) + coalesce(adult, 0) +coalesce( elderly,0 )) Cases 
                              from  tb
                                group by 
                                country, year 
                              having 
                                sum( coalesce(child, 0) + coalesce(adult, 0) +coalesce( elderly,0 ))  > 0")
kable(head ( tbCases) ) 

```


```{r warning=FALSE}
populationWithTbCases <- left_join(x= tbCases, y=population)
kable(head(populationWithTbCases))
```

```{r}
populationWithTbCases$country <- as.factor(populationWithTbCases$country)
populationWithTbCases$rate <- populationWithTbCases$Cases / populationWithTbCases$population
kable(head(populationWithTbCases))
```


You may wnat to visualize this data to see where TB hotspots are, or where TB is on the rise so you could intervene and stop the spread.  
