---
title: "CUNY-DATA-607-Week7-JSON-XML-HTML"
author: "Raphael Nash"
date: "10/8/2016"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library("rvest")
library("knitr")
library("jsonlite")
library("RCurl")
```

#Overview

The purpose of this assignment is to explore the difference between JSON, XML and HTML files.  Create a file of each type to store information on 3 books.  Then parse the files into a data frame.  Are all three data frames the same?

#Parse XML
```{r}

#Create empty data frame
xml_books_df <- data.frame(id = character(),
                           title = character(),
                          authors = I ( list() ) ,
                           price = numeric()
                           )
```
   
XML file looks like: 
<books>
  <book>
  .
  .
  .
  </book>
<book>
```{r}                        
#read file from github                           
xml_file <- getURL ("https://raw.githubusercontent.com/RaphaelNash/CUNY-DATA-607/master/Week7-HTMLKMLJSON/books.xml") %>% 
  read_xml()

#loop through all of the "book" nodes in the file
for ( i in 1:length(xml_nodes(xml_file, "book"))) {

  book_node <-html_nodes(xml_file, "book")[i]
  
  # get the ID
  id<- html_nodes(book_node, "id") %>% 
    xml_text()
  
  # get the title 
  title<- html_nodes(book_node, "title") %>% 
    xml_text()
    
  #get the authors list   
  authors_node <- html_nodes(book_node, "authors") 
  
  authors_list <- lapply(authors_node, function(x)  html_nodes(authors_node, "author") %>% 
    xml_text() ) %>% 
    I()
    
  # get the price 
  price <- html_nodes(book_node, "price") %>% 
    xml_text()
 
 #create a row in a temporary data frame
 temp_df <- data.frame(
   id = id,
   title = title,
   price = price,
   authors = authors_list
 )

#add the temp row to the main data frame
xml_books_df <- rbind(xml_books_df, temp_df)
}
    

kable(xml_books_df)
```

#Parse HTML
```{r}

#Read file from git hub and extract the table
html_nodes  <- getURL ("https://raw.githubusercontent.com/RaphaelNash/CUNY-DATA-607/master/Week7-HTMLKMLJSON/books.html") %>% 
  read_html() %>% 
  html_nodes( xpath="//table[1]")

table <- html_nodes[1] 

#Convert to cells to a data frame
 html_books_df <- html_table(table, fill=TRUE, header = TRUE)[[1]]

kable(html_books_df)
```

#Parse JSON
```{r}

#Read JSON file from github
json_file <-  getURL ("https://raw.githubusercontent.com/RaphaelNash/CUNY-DATA-607/master/Week7-HTMLKMLJSON/books.json")

#Convert JSON to a data frame
json_books_df <- fromJSON(json_file)

kable(json_books_df)
```

#Looks at Structure of Data.Frames
```{r}
str(xml_books_df)

str(html_books_df)

str(json_books_df)
```

#Conclusions
The data frames created from the XML and JSON files are identical.  The only difference between the HTML file and the other files is that the authors column is a list object in the XML and JSON files while in the HTML file the authors field is just a text comma seperated list.  
