---
title: "Lab4-ggplot"
author: "Raphael Nash"
date: "September 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Data

1. Load the diamonds dataset, which is included with the ggplot2 package. Identify which variables in the diamond set are numeric, and which are categorical (factors).

```{r}
library(ggplot2)
library(knitr)
library(DT)
library(plotly)
head(diamonds)
str(diamonds)
```

Varaible | Variable Type
---------|-------------
cut      |Factor
carat    |Factor
color    |Factor
clarity  |Factor
depth    |Numeric
table    |Numeric
price    |Numeric
x        |Numeric
y        |Numeric
x        |Numeric

#Summary Stats
2. Generate summary level descriptive statistics: Show the mean, median, 25th and 75th quartiles, min, and max for each of the applicable variables in diamonds.

```{r}
summary(diamonds)
```

3. Determine the frequency for each of the diamond colors.
```{r}
table(diamonds$color)
ggplot(diamonds, aes(color)) + geom_bar()
```

4. Determine the frequency for each of the diamond cuts, by color
```{r}
table(diamonds$cut, diamonds$color)
ggplot(diamonds, aes( cut, fill=color )) + geom_bar()
```

5. Create a graph for a single numeric variable.
```{r}

ggplot(diamonds, aes(price)) + geom_histogram(binwidth = 100)
ggplot(diamonds, aes(price)) + geom_histogram(binwidth = 100)+facet_grid(cut ~ .)
 ggplot(diamonds, aes(price)) + geom_histogram(binwidth = 500)+facet_grid(cut ~ clarity) +ylim(0,600)
```


```{r}
ggplot(diamonds, aes(x=price, y=carat, color = clarity)) +geom_point()
```

```{r}
p <- ggplot(diamonds, aes(x=price, y=carat, color = clarity)) +geom_point()
ggplotly()
```