---
title: "Week3-Assignment-RegEx"
author: "Raphael Nash"
date: "September 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
```



```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"   
```

3.Copy the intorductory vector name stores the extracted names 

```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
```

3.1. Use the tools of this chapter to re-arrange the vector so that all of the elements conform to the standard first_name last_name

```{r}
first_last <- function(x){
      
  if ( str_detect(x,",")){
    firstAndLast <- unlist(str_extract_all(x, "[[:alpha:]]{3,}")) 
    currentName <- paste(trimws(firstAndLast[2]),  firstAndLast[1])
  }  else if (str_detect(x, "^[:alpha:]{2,3}\\.") ) {
      firstAndLast <- unlist(str_extract_all(x, "[[:alpha:]]{4,}"))
      currentName <- paste(firstAndLast[1],  firstAndLast[2])
    } else {
      currentName <- x
      
    }
  currentName
}

standardNameFormat <- unname(sapply( name, first_last ))
standardNameFormat
```

3.2 Construct A logical Vector indicating whether a character has a title (Rev. or Dr.)

```{r}

hasTitle <- str_detect(name, "^[:alpha:]{2,3}\\.") 

name
hasTitle

```

3.2 Construct A logical Vector indicating whether a character has a second name

```{r}

hasSecondName <- str_detect(name, "[:space:][:alpha:]{1}\\.[:space:][:alpha:]+") 

name
hasSecondName

```

4. Describe the types of strings that confom to the following regular expresstions and construct and example that is matched by the regular expression.


4.1 [0-9]+\\$

One or more digits followed by a Dollar Sign
```{r}
 str_detect ("1234$", "[0-9]+\\$")
```

4.2 \\b[a-z]{1,4}

The begining of a line starts with 1 to 4 occurrences of a lower case letter
```{r}
str_detect ("a", "\\b[a-z]{1,4}")
str_detect ("ab", "\\b[a-z]{1,4}")
str_detect ("abc", "\\b[a-z]{1,4}")
str_detect ("abcd", "\\b[a-z]{1,4}")
```

4.3 .*?\\.txt$

Anything with an optional space and then the line ends with ".txt"

```{r}
str_detect("abc .txt", ".*?\\.txt$")
str_detect("abc.txt", ".*?\\.txt$")
```

4.4 \\d{2/\\d{2}/\\d{4}}

Standard date format mm/dd/yyyy


```{r}
str_detect("01/01/2016", "\\d{2}/\\d{2}/\\d{4}")
```

4.5  <(.+?)>.+</\\1>

This checks to see if you have HTML/XML tags that are properly closed with optional text between them. 

```{r}
str_detect("<foo>bar</foo>", "<(.+?)>.+</\\1>")
```